---
title: "HS614 HW#1"
author: "Tiffany Chua"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HS614 Homework #1
```{r Working Directory}
getwd()
```

```{r Libraries}
library("nycflights13")
library("dplyr")
library("tidyverse")
```

## R Exercise

1. Create a function  called "count_func" that accepts two arguments, a string and a vector of string. It returns the count of the number of occurrences of the string in the input vector.
```{r 1. count_func}
count_func <- function(string, v){
  count = 0
  for (x in v) {
    if (x == string) {
      count = count + 1
    }
  }
  return (count)
}

#Example:
x = c("ML", "ML", "data", "data2")
count_func("ML", x)
```

2. Write a function called "get_min" that uses if, else if, and else statements to print in a numeric vector with 3 elements.
```{r 2. get_min}
get_min <- function(v) {
  if ((x[1] < x[2]) & (x[1] < x[3])) {
    print (x[1])
  }
  else if (x[2] < x[3]) {
    print(x[2])
  }
  else {
    print(x[3])
  }
}

#Example:
v = c(-110, 20, -30)
get_min(v) 
# Return: Error in if (x[1] < x[2] & x[1] < x[3]) { : 
#  missing value where TRUE/FALSE needed
```

3. Create a function called "get_even_odd" that accepts an integer and will tell you if it is odd or even.
```{r 3. get_even_odd}
get_even_odd <- function(x){
  if (x%%2 == 0) {
    print("Even")
  }
  
  else {
    print("Odd")
  }
}

#Example"
get_even_odd(-3)
get_even_odd(6)
```

4. Recreate the following dataframe (see assignment) by creating vectors and using the data.frame function
```{r 4. dataframe}
A <- c("a", "b", "c")
B <- c(1, 2, 3)
C <- c(TRUE, FALSE, TRUE)

data.frame(A, B, C)
```

5. Print top rows (head) of the flights dataset and answer the following questions:
```{r 5. head}
head(flights)
```

  * a. How many rows and columns are in the dataset?
```{r 5a}
nrow(flights)
ncol(flights)
```
  
  * b. How many years do we have data for in the dataset?
```{r 5b}
unique(flights$year)
# 1 year: 2013
```

  * c. What are the number of flights in each month?
```{r 5c}
table(flights$month)
```
  
  * d. Does it contain any missing values?
```{r 5d}
sum(is.na(flights))
nrow(flights)

#There are 46595 rows with missing values
```
  
  * e. Select the rows that are at the last day of the months (30 or 31). Print top 6 rows (head).
```{r 5e}
head(flights %>% filter(day == 30 | day == 31))
```
  
  * f. What is the average of "dep_time"?
```{r 5f}
summary(flights)
# Mean = 1349
```
  
  * g. Select only columns of "year", "month", "day", "carrier", "origin", "dest". Print the head.
```{r 5g}
head(flights[ ,
         c("year", "month", "day", "carrier", "origin", "dest")])
```
  
  