---
title: "HS614 HW#2"
author: "Tiffany Chua"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries}
library(Amelia)
library(ggplot2)
library("GGally")
library(caTools)
library(caret)
library(class)
library("e1071")
library(psych)
library(corrplot)
library("ISLR")
```

##HW #2:
For this HW do more exploratory data analysis (EDA) like adding some graphs, correlation btw variables. After that apply all the classification models we have learned so far. 

```{r Dataset}
# https://archive.ics.uci.edu/ml/datasets/breast+cancer+wisconsin+(original)
url = "https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/breast-cancer-wisconsin.data"
names = c('id', 'clump_thickness', 'uniform_cell_size', 'uniform_cell_shape',
       'marginal_adhesion', 'single_epithelial_size', 'bare_nuclei',
       'bland_chromatin', 'normal_nucleoli', 'mitoses', 'class')
df = read.csv(url, col.names=names)

df
```

Attribute Domain
1. Sample code number id number
2. Clump Thickness 1 - 10
3. Uniformity of Cell Size 1 - 10
4. Uniformity of Cell Shape 1 - 10
5. Marginal Adhesion 1 - 10
6. Single Epithelial Cell Size 1 - 10
7. Bare Nuclei 1 - 10
8. Bland Chromatin 1 - 10
9. Normal Nucleoli 1 - 10
  A. Mitoses 1 - 10
  B. Class: (2 for benign, 4 for malignant)
  
  
```{r}
str(df)
```
  
```{r}
#Change bare_nuclei to int
df$bare_nuclei <- as.integer(df$bare_nuclei)

#Change class to factor
df$class = factor(df$class,
                  levels = c(2, 4),
                  label = c(0, 1))

str(df)
```
  
```{r}
nrow(df)
ncol(df)
```

```{r}
table(df$bare_nuclei)
```

```{r}
df = subset(df,
            df$bare_nuclei != "?")

table(df$bare_nuclei)
```

```{r}
str(df)

sum(is.na(df))
```

```{r}
df = df[ , -1]
str(df)
```
  
```{r}
summary(df)
```

```{r}
missmap(df,
        main = "Missing Map",
        col = c("blue",
                "red"))
```

```{r}
ggpairs(df)
```

```{r}
#Check how many in each class
table(df$class)
```

