---
title: "HS614 HW#3"
author: "Tiffany Chua"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Working Directory}
getwd()
```

# HS614: Homework #3
For this HW use the diabetes dataset.

Exploratory Data Analysis:
1. Explore the data
2. How many rows and columns?
3. Any NAs? If yes, which columns?
4. Any strange values? If yes, what to do?
5. Any correlation between features?
6. Statistics of the columns and some exploratory plots (bar plot, histogram, boxplot ...)

Build at least 4 different predictive models that can predict (classify) Diabetic patients. Compare the performance of the models.

Column definitions of this dataset are given in table 1 of this paper: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8306487/

***

```{r Libraries}
library(Amelia)
library(ggplot2)
library(GGally)
library(caTools)
library(caret)
library(class)
library(e1071)
library(psych)
library(corrplot)
library(ISLR)
library(cluster)
library(datasets)
library(pastecs)
library(psych)
library(corrplot)
library(stats)
library(Hmisc)
library(mice)
library(VIM)
library(dplyr)
library(hrbrthemes)
library(tidyverse)
library(factoextra)
library(pROC)
library(rpart)
library(rpart.plot)
library(randomForest)
library(animation)
```

```{r Read in Dataset}
diab <- read.csv("./diabetes.csv")
head(diab)
str(diab)
```

## Variable Definitions
1. Pregnancy - Frequency of pregnancy
2. Glucose - Concentration of plasma glucose (mg/dL)
3. BP - Diastolic blood pressure (mmHg)
4. Skin - Tricep skinfold thickness (mm)
5. Insulin - Two-hour serum insulin (mu U/ml)
6. BMI - Body mass index (kg/m2)
7. Pedigree - A pedigree function for diabetes
8. Age - Age (log(years))

```{r Outcome Factor}
diab$Outcome = factor(diab$Outcome)
str(diab)

table(diab$Outcome)
```

## Rows & Columns
```{r}
nrow(diab)
ncol(diab)
```

## Missing & Strange Variables
```{r Missing/NA}
sum(is.na(diab))
missmap(diab)

#Note: While there are no NA variables, missing data points seem to have been replaced with a value of 0 in columns
```
Note: While there are no NA variables, missing data points seem to have been replaced with a value of 0 in columns.

```{r Rows with Value 0 per Column}
sum(diab$Glucose == 0)
sum(diab$BloodPressure == 0)
sum(diab$SkinThickness == 0)
sum(diab$Insulin == 0)
sum(diab$BMI == 0)
sum(diab$DiabetesPedigreeFunction == 0)
sum(diab$Age == 0)

#Strange Values:
#The following columns returned values of 0 Glucose, BloodPressure, SkinThickness, Insulin, and BMI; It is likely that missing/NA variables for these columns were replaced with 0, as 0 is a highly unlikely and unrealistic observation for these features
```

