---
title: "HS614 HW#3"
author: "Tiffany Chua"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Working Directory}
getwd()
```

# HS614: Homework #3
For this HW use the diabetes dataset.

Exploratory Data Analysis:
1. Explore the data
2. How many rows and columns?
3. Any NAs? If yes, which columns?
4. Any strange values? If yes, what to do?
5. Any correlation between features?
6. Statistics of the columns and some exploratory plots (bar plot, histogram, boxplot ...)

Build at least 4 different predictive models that can predict (classify) Diabetic patients. Compare the performance of the models.

Column definitions of this dataset are given in table 1 of this paper: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8306487/

***

```{r Libraries}
library(Amelia)
library(ggplot2)
library(GGally)
library(caTools)
library(caret)
library(class)
library(e1071)
library(psych)
library(corrplot)
library(ISLR)
library(cluster)
library(datasets)
library(pastecs)
library(psych)
library(corrplot)
library(stats)
library(Hmisc)
library(mice)
library(VIM)
library(dplyr)
library(hrbrthemes)
library(tidyverse)
library(factoextra)
library(pROC)
library(rpart)
library(rpart.plot)
library(randomForest)
library(animation)
```

```{r Read in Dataset}
diab <- read.csv("./diabetes.csv")
head(diab)
str(diab)
```

